---
title: "Elasticity and Linear Models"
author: "Juan Malaver"
date: "9/13/2021"
output:
  ioslides_presentation:
    df_print: paged
    widescreen: true
    logo: logo_square.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(readr)
library(ggplot2)
library(broom)

# importing data
oj <- read_csv("data/oj.csv")
```

## Business Problem

We want to model and predict a *response* variable ($y=\text{sales}$) as a 
function of *inputs* ($x=\text{price}$).

$$\mathbb{E}[y|\boldsymbol{x}]=\boldsymbol{x'\beta}=\beta_0+x_1\beta_1+\ldots+x_p\beta_p$$

We will use *price elasticity of demand* ($\beta$) to understand how changes in
`price` impact changes in `sales`.

$$\text{log}(\texttt{sales})=\alpha+\beta\text{log}(\texttt{price})+\epsilon$$

## Orange Juice Data

```{r echo=FALSE}
head(oj, n = 100)
```

## Prices by Brand {.flexbox .vcenter}

```{r echo=FALSE}
ggplot(data = oj, mapping = aes(x = brand, y = log(price), col = brand)) +
  geom_boxplot()
```

## Prices and Sales by Brand {.flexbox .vcenter}

```{r echo=FALSE}
ggplot(data = oj, mapping = aes(x = log(price), y = log(sales), col = brand)) +
  geom_point(alpha = 0.25)
```

## Interpreting Price Elasticity

`sales` increase by $\beta\%$ for every $1\%$ increase in `price`.

$$\frac{dy}{dx}=\frac{\beta}{x}e^{\alpha+\beta\text{log}(x)+\epsilon}\Rightarrow\beta=\frac{dy/y}{dx/x}$$

`sales` and `price` must be strictly non-negative and we assume their conditional *distribution* follows
a Normal distribution.

$$y|\boldsymbol{x}\sim \mathcal{N}(\boldsymbol{x'\beta},\sigma^2)$$

## Price Elasticity by Independent Brands

```{r echo=FALSE}
mod <- glm(log(sales) ~ log(price) + brand, data=oj)
tidy(mod)
```

## Random Intercepts{.flexbox .vcenter}

```{r echo=FALSE}
ggplot(
  data = cbind(oj, pred = predict(mod)),
  mapping = aes(
    x = log(price),
    y = log(sales),
    color = brand
  )
) + geom_point(col = "gray") +
  geom_line(mapping = aes(y = pred), size = 1)
```

## Price Elasticity by Dependent Brands

```{r echo=FALSE}
mod <- glm(log(sales) ~ log(price)*brand, data=oj)
tidy(mod)
```

## Random Slopes Random Intercepts{.flexbox .vcenter}

```{r echo=FALSE}
ggplot(
  data = cbind(oj, pred = predict(mod)),
  mapping = aes(
    x = log(price),
    y = log(sales),
    color = brand
  )
) + geom_point(col = "gray") +
  geom_line(mapping = aes(y = pred), size = 1)
```

## References

Taddy, M. (2019). *Business data science: Combining machine learning and economics to optimize, automate, and accelerate business decisions* (1 Edition). McGraw-Hill Education.

Varian, H. R. (2014). *Intermediate microeconomics: A modern approach* (Ninth edition, international student edition). W.W. Norton & Company.
